---
title: "Transcript Annotation Database for *Glycine max*"
author: "Kabilan S"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Transcript Annotation Database for Glycine max}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

**TxDb.Gmax.JGI.Gmv4** is a Bioconductor-compatible transcript annotation
package for *Glycine max* generated from the official **JGI/Phytozome GFF3
annotation** of the **Glycine max Wm82.a4.v1** reference genome.

The package provides genomic coordinates and transcript models, including genes,
transcripts, exons, coding sequences (CDS), and promoter regions. It is intended
for downstream genomic analyses using Bioconductor packages such as
**GenomicFeatures**, **GenomicRanges**, **Signac**, **ArchR**, and
**ChIPseeker**.

The TxDb object itself is not shipped inside the package; it is hosted on
**AnnotationHub** and downloaded on first use.

```{r eval=FALSE}
# Installation
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install(c(
    "AnnotationHub",
    "GenomicFeatures",
    "GenomicRanges"
))
```

## Loading packages
```{r loading, message=FALSE}
library(AnnotationHub)
library(GenomicFeatures)
```

```{r downloading}
# Downloading the TxDb file from AnnotationHub. Run
# query(ah, "TxDb.Gmax.JGI.Gmv4") to look up the AH accession.
TxDb.Gmax.JGI.Gmv4 <- tryCatch({
    ah <- AnnotationHub()
    ah[["AHxxxxxx"]]
}, error = function(e) NULL)
```

```{r gate, include=FALSE}
## The remaining chunks need the downloaded resource; skip them when the
## build machine has no network access or the resource is unavailable.
knitr::opts_chunk$set(eval = !is.null(TxDb.Gmax.JGI.Gmv4))
```

# TxDb file summary

```{r summary}
TxDb.Gmax.JGI.Gmv4
```

# Chromosome information

```{r chromosome}
seqinfo(TxDb.Gmax.JGI.Gmv4)
```

# Available chromosomes

```{r avl_chromosomes}
seqlevels(TxDb.Gmax.JGI.Gmv4)
```
The 20 nuclear chromosomes of the Glycine max Wm82.a4.v1 assembly. These
contain the vast majority of annotated genes.

# Gene annotations

Retrieve all annotated genes.

```{r gene_annotations}
genes(TxDb.Gmax.JGI.Gmv4)
```

# Transcript annotations

```{r transcripts_annotations}
head(transcripts(TxDb.Gmax.JGI.Gmv4))
```

# Exon annotations

```{r exon_annotations}
head(exons(TxDb.Gmax.JGI.Gmv4))
```

# Coding sequence (CDS) annotations

```{r cds}
head(cds(TxDb.Gmax.JGI.Gmv4))
```

# Promoter regions

Extract promoter regions 2 kb upstream and 200 bp downstream of transcription
start sites.

```{r promoters}
head(promoters(
    TxDb.Gmax.JGI.Gmv4,
    upstream = 2000,
    downstream = 200
))
```

# Transcripts grouped by gene

```{r transcripts_gene}
tx_by_gene <- GenomicFeatures::transcriptsBy(TxDb.Gmax.JGI.Gmv4, by = "gene")
tx_by_gene
```

# Exons grouped by transcript

```{r exons_transcript}
exonsBy(TxDb.Gmax.JGI.Gmv4, by = "tx")
```

# Session information

```{r sessioninfo, eval=TRUE}
sessionInfo()
```
