---
title: "Transcript Annotation Database for *Glycine max*"
author: "Kabilan S"
date: "2026-08-15"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Transcript Annotation Database for Glycine max}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

**TxDb.Gmax.JGI.Gmv4** is a Bioconductor-compatible transcript annotation
package for *Glycine max* generated from the official **JGI/Phytozome GFF3
annotation** based on the **Glycine max Wm82.a4.v1** reference genome.

The package provides genomic coordinates and transcript models, including
genes, transcripts, exons, coding sequences (CDS), and promoter regions. It is
intended for downstream genomic analyses using Bioconductor packages such as
**GenomicFeatures**, **GenomicRanges**, **Signac**, **ArchR**, and
**ChIPseeker**.

The TxDb object is retrieved from **AnnotationHub** using its corresponding
AnnotationHub record.

# Installation

Install the package and its dependencies from Bioconductor with:

```{r install, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install(c(
    "TxDb.Gmax.JGI.Gmv4",
    "AnnotationHub",
    "GenomicFeatures",
    "GenomicRanges"
))
```

# Loading the package

```{r loading, message=FALSE, warning=FALSE}
library(TxDb.Gmax.JGI.Gmv4)
library(AnnotationHub)
library(GenomicFeatures)
library(GenomicRanges)
```

# Retrieve the TxDb object from AnnotationHub

The TxDb object is available through `AnnotationHub`:

```{r annotationhub, message=FALSE, warning=FALSE}
ah <- AnnotationHub::AnnotationHub()
TxDb.Gmax.JGI.Gmv4 <- ah[["AH122288"]]
```

Check the class of the retrieved object:

```{r class}
class(TxDb.Gmax.JGI.Gmv4)
```

# Package summary

Display the transcript database:

```{r summary}
TxDb.Gmax.JGI.Gmv4
```

A summary can also be obtained with:

```{r summary-function}
summary(TxDb.Gmax.JGI.Gmv4)
```

# Chromosome information

The chromosome and sequence information can be examined with `seqinfo()`:

```{r chromosome}
seqinfo(TxDb.Gmax.JGI.Gmv4)
```

# Available chromosomes

The available sequence levels can be obtained with:

```{r avl-chromosomes}
seqlevels(TxDb.Gmax.JGI.Gmv4)
```

The Wm82.a4.v1 assembly contains 20 nuclear chromosomes, which contain the
majority of the annotated genes.

# Gene annotations

Gene ranges can be retrieved with `genes()`:

```{r gene-annotations}
gene_annotations <- genes(TxDb.Gmax.JGI.Gmv4)
gene_annotations
```

# Transcript annotations

Transcript ranges can be retrieved with `transcripts()`:

```{r transcript-annotations}
transcript_annotations <- transcripts(TxDb.Gmax.JGI.Gmv4)
head(transcript_annotations)
```

# Exon annotations

Exon ranges can be retrieved with `exons()`:

```{r exon-annotations}
exon_annotations <- exons(TxDb.Gmax.JGI.Gmv4)
head(exon_annotations)
```

# Coding sequence (CDS) annotations

Coding sequence ranges can be retrieved with `cds()`:

```{r cds}
cds_annotations <- cds(TxDb.Gmax.JGI.Gmv4)
head(cds_annotations)
```

# Promoter regions

Promoter regions can be generated using 2 kb upstream and 200 bp downstream
of transcription start sites:

```{r promoters}
promoter_annotations <- promoters(
    TxDb.Gmax.JGI.Gmv4,
    upstream = 2000,
    downstream = 200
)
head(promoter_annotations)
```

# Transcripts grouped by gene

Transcripts can be grouped by gene using `transcriptsBy()`:

```{r transcripts-by-gene}
tx_by_gene <- GenomicFeatures::transcriptsBy(
    TxDb.Gmax.JGI.Gmv4,
    by = "gene"
)
tx_by_gene
```

# Exons grouped by transcript

Exons can be grouped by transcript using `exonsBy()`:

```{r exons-by-transcript}
exons_by_transcript <- GenomicFeatures::exonsBy(
    TxDb.Gmax.JGI.Gmv4,
    by = "tx"
)
exons_by_transcript
```

# Session information

```{r sessioninfo}
sessionInfo()
```
