---
title: "Transcript Annotation Database for *Glycine max*"
author: "Kabilan S"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Transcript Annotation Database for Glycine max}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

**TxDb.Gmax.JGI.Gmv4** is a Bioconductor-compatible transcript annotation
package for *Glycine max* generated from the official **JGI/Phytozome GFF3 annotation** GFF3 annotation based on the **Glycine max Wm82.a4.v1** reference genome.

The package provides genomic coordinates and transcript models, including genes,
transcripts, exons, coding sequences (CDS), and promoter regions. It is intended
for downstream genomic analyses using Bioconductor packages such as
**GenomicFeatures**, **GenomicRanges**, **Signac**, **ArchR**, and **ChIPseeker**.

The TxDb object itself is not shipped inside the package; it is hosted on
**AnnotationHub** and downloaded on first use.

```{r eval=FALSE}
# Install required Bioconductor packages
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(c(
    "GenomicFeatures",
    "GenomicRanges"
))
```

# Loading the package

```{r loading}
library(TxDb.Gmax.JGI.Gmv4)
library(GenomicFeatures)
```

# Package summary

```{r summary}
TxDb.Gmax.JGI.Gmv4
```

# Chromosome information

```{r chromosome}
seqinfo(TxDb.Gmax.JGI.Gmv4)
```

# Available chromosomes

```{r avl_chromosomes}
seqlevels(TxDb.Gmax.JGI.Gmv4)
```
The 20 nuclear chromosomes of the Glycine max Wm82.a4.v1 assembly. These contain the vast majority of annotated genes. 

# Gene annotations

Retrieve all annotated genes.

```{r gene_annotations} 
genes(TxDb.Gmax.JGI.Gmv4)
```

# Transcript annotations

```{r transcripts_annotations}
head(transcripts(TxDb.Gmax.JGI.Gmv4))
```

# Exon annotations

```{r exon_annotations}
head(exons(TxDb.Gmax.JGI.Gmv4))
```

# Coding sequence (CDS) annotations

```{r cds}
head(cds(TxDb.Gmax.JGI.Gmv4))
```

# Promoter regions

Extract promoter regions 2 kb upstream and 200 bp downstream of transcription start sites.

```{r promoters}
head(promoters(
    TxDb.Gmax.JGI.Gmv4,
    upstream = 2000,
    downstream = 200
))
```

# Transcripts grouped by gene

```{r transcripts_gene}
tx_by_gene <- GenomicFeatures::transcriptsBy(TxDb.Gmax.JGI.Gmv4, by = "gene")
tx_by_gene
```

# Exons grouped by transcript

```{r exons_transcript}
exonsBy(TxDb.Gmax.JGI.Gmv4, by = "tx")
```

# Session information

```{r sessioninfo, eval=TRUE}
sessionInfo()
```
